---
title: "DATA 555 Interactive Visulizations"
author: "Xintian Song"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, message=FALSE}
library(flexdashboard)
library(ggplot2)
library(shiny)
library(plotly)
library(DT)
# library(manipulate)

library(tidyverse)
library(magrittr)
library(here)
```

```{r, include=FALSE}
# knitr::opts_knit$set(root.dir = here::here())
# setwd(knitr::opts_knit$get("root.dir"))
```

```{r, include=FALSE}
options(shiny.sanitize.errors = FALSE)

# shiny::addResourcePath("data", rprojroot::find_rstudio_root_file("final project", "data"))

baby <- readRDS("data/baby.rds")
mother <- readRDS("data/mother.rds")
tr <- readRDS("data/tr.rds")
hb <- readRDS("data/hb.rds")

# DATA_FOLDER <- "final project/data"
# baby <- readRDS(here::here("data", "baby.rds"))
# mother <- readRDS(here::here("data",  "mother.rds"))
# tr <- readRDS(here::here("data",  "tr.rds"))
# hb <- readRDS(here::here("data", "hb.rds"))
```

```{r}
baby2 <- baby %>%
  mutate(center = case_when(
    startsWith(as.character(id), "11") ~ "Emory",
    startsWith(as.character(id), "12") ~ "Grady",
    startsWith(as.character(id), "13") ~ "Northside",
    startsWith(as.character(id), "14") ~ "Hopkins",
    TRUE ~ "Other"  
  )) %>%
  select(id, center, LBWIDOB, gender, race, gestagewk, gestageday, birthweight002, 
         apgar1min, apgar5min) %>%
  rename(baby_id = id, birthweight=birthweight002) %>%
  mutate(gestageweek = (7*gestagewk + gestageday)/7 ) %>%
  select(-c(gestagewk, gestageday))

tr2 <- tr %>%
  left_join(baby2, by = "baby_id") %>%
  left_join(mother, by = "baby_id") %>%
  select(-moc_id) %>%
  
  mutate(gestageweek = case_when(
    is.na(gestageweek) ~ NA_character_,
    gestageweek < 28 ~ "Extreme Preterm",    
    gestageweek < 32 ~ "Very Preterm",
    TRUE ~ "Moderate Preterm"
  )) %>%
  
  mutate(MOC_race003 = dplyr::recode(MOC_race003,
    `1` = "Black",
    `2` = "American Indian",
    `3` = "White",
    `4` = "Native Hawaiian or Other Pacific Islander",
    `5` = "Asian",
    `6` = "More than one race",
    `7` = "Other",
    `99` = NA_character_
  )) %>%
  
  mutate(race = dplyr::recode(race,
    `1` = "Black",
    `2` = "American Indian",
    `3` = "White",
    `4` = "Native Hawaiian or Other Pacific Islander",
    `5` = "Asian",
    `6` = "More than one race",
    `7` = "Other",
    `99` = NA_character_
  )) %>%
  
  
  mutate(
    MultipleBirth004 = ifelse(MultipleBirth004 == 99, NA, MultipleBirth004),
    RHgroup004 = ifelse(RHgroup004 == 99, NA, RHgroup004)
  ) %>%
  
  mutate(gender = dplyr::recode(gender,
    `1` = "Female",
    `2` = "Male"
  )) %>%
  
  mutate(across(-c(1:5), as.factor))


tr2 %<>% rename(
  `Mother Race Group` = MOC_race003,
  `Mother Insurance Group` = Insurance003,
  `Mother Blood Rh Group` = RHgroup004,
  `Infant Gender Group` = gender,
  `Infant Race Group` = race, 
  `Does Infant belong to Multiple Birth?` = MultipleBirth004,
  `Infant Gestational age` = gestageweek
)
```

# Dataset Description 

- The dataset used in the project is confidential.

- We included 284 preterm infants in the dataset, with 832 transfusion records and 3334 hemoglobin level measurements. 

- The data was collected with a secondary, prospective, multicenter observational cohort study for preterm infants who were born in 3 hospitals between 2017-01-30 and 2024-06-09. 


# RBC Transfusion Needs

## Column {.sidebar data-width="300"}

```{r}
# x_choice <- menu(x_options, title = "Select an X-axis variable:")
# x_axis <- x_options[x_choice]

x_options <- c("Mother Race Group", "Mother Insurance Group", "Mother Blood Rh Group", 
               "Infant Gender Group", "Infant Race Group", "Does Infant belong to Multiple Birth?",
               "Infant Gestational age")

selectInput(inputId = "x_axis", 
            label = h4("Select X-Axis:"), 
            choices = x_options, selected = "MOC_race003")


selectInput("center", h4("Select Center:"), 
              choices = c("All", "Emory", "Grady", "Northside", "Hopkins"),
              selected = "All")
```


## Column {data-width="350"}
- **My GitHub Repo**
  - https://github.com/MOI13/ILE_dashboard

- **Project Objective 1**
  - Identify the preterm infant subgroups most in need of RBC transfusions.

- **Key Takeaways**
  - Infants born to minority mothers, those covered by Medicaid insurance, female infants, infants from multiple births (twins/triplets), and those with a gestational age between 32 and 37 weeks tend to receive the highest volume of RBC transfusions per transfusion event.
  - This pattern remains consistent across different centers.

```{r}
renderPlotly({
  
  tr3 <- if (input$center == "All") {
    tr2
  } else {
    tr2 %>% filter(center == input$center)
  }
  
  x_axis <- input$x_axis

  plot_data <- tr3 %>%
    select(all_of(x_axis), rbcVolumeTransfused031, baby_id) %>%
    drop_na() %>%
  mutate(!!x_axis := case_when(
    .data[[x_axis]] == 1 ~ "Yes",
    .data[[x_axis]] == 0 ~ "No",
    TRUE ~ as.character(.data[[x_axis]])
  )) %>%
    group_by(!!sym(x_axis)) %>%
    mutate(medianrbc = median(rbcVolumeTransfused031, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(fct_in = fct_reorder(as.factor(.data[[x_axis]]), medianrbc, .desc = TRUE))
  

  p <- ggplot(plot_data, aes(x = fct_in, y = rbcVolumeTransfused031)) +  
    geom_boxplot(outlier.shape = NA, fill = "lightblue", alpha = 0.5) +
    geom_jitter(aes(text = paste("\n", x_axis, ":", .data[[x_axis]], 
                                 "\nInfant ID:", baby_id,
                                 "\nRed Blood Cell Volume:", rbcVolumeTransfused031)), 
                width = 0.2, color = "darkred", alpha = 0.6) +
    labs(x = x_axis, y = "RBC Transfusion Amount (mL)", 
         title = "Red Blood Cell Transfusion Distribution") +
    theme_bw() +
     theme(panel.grid.major = element_blank(),  
          panel.grid.minor = element_blank()) 

  ggplotly(p, tooltip = "text") %>% 
    layout(
      height = 600,  
      annotations = list(
        list(
          x = 0.16, y = 1, 
          text = "Gestational Age Classification:\nExtreme Preterm: <28 weeks\nVery Preterm: 28-31 weeks\nModerate Preterm: 32+ weeks", 
          showarrow = FALSE, xref = 'paper', yref = 'paper', 
          xanchor = 'right', yanchor = 'auto', xshift = 0, yshift = 0, 
          font = list(size = 11, color = "grey")
        )
      )
    )
})
```

# Hemoglobin Level & Body Weight Trend

```{r}
tr4 <- tr2 %>%
  mutate(tran_day = as.numeric(interval(as.Date(LBWIDOB), as.Date(DateTransfusion031)) / days(1))) %>%
  left_join(hb %>%
              arrange(baby_id, infant_day) %>%  
              group_by(baby_id, infant_day) %>%
              slice(1) %>%
              ungroup() %>%
              dplyr::select(baby_id, recent_hb = HB, infant_day), 
          join_by(baby_id, closest(tran_day >= infant_day))) %>%
  rename(`Body Weight(g)` = BodyWeight031,
         `Hemoglobin Level(g/dl)` = recent_hb)

```

## Column {.sidebar data-width="300"}

```{r}
x_options <- c("Body Weight(g)", "Hemoglobin Level(g/dl)")

selectInput(inputId = "x_axis_2", 
            label = h4("Select X-Axis:"), 
            choices = x_options, selected = "Body Weight(g)")


selectInput("center_2", h4("Select Center:"), 
              choices = c("All", "Emory", "Grady", "Northside", "Hopkins"),
              selected = "All")
```


## Column {data-width="350"}
- **My GitHub Repo**
  - https://github.com/MOI13/ILE_dashboard

- **Project Objective 2**
  - Assess how RBC transfusions affect the health outcomes of preterm infants.

- **Key Takeaways**
  - Overall, preterm infants’ body weight tends to increase with each transfusion event over time. Infants born at Grady and Northside hospitals’ NICUs experience the most significant weight gain.
  - In general, preterm infants’ hemoglobin levels decrease with each transfusion event and stabilize around 3 weeks after birth. Infants born at Emory have the highest baseline hemoglobin levels, while those born at Grady have the lowest baseline levels on average, indicating poorer health conditions at birth. However, after multiple RBC transfusion events, hemoglobin levels in all infants stabilize and remain relatively similar after 3 weeks.


```{r}
renderPlotly({
  
  tr5 <- if (input$center_2 == "All") {
    tr4
  } else {
    tr4 %>% filter(center == input$center_2)
  }
  
  x_axis <- input$x_axis_2

  plot_data <- tr5 %>%
    select(all_of(x_axis), tran_day, baby_id) %>%
    drop_na() %>%
    mutate(across(everything(), as.numeric))
  

  p <- ggplot(plot_data, aes(x = tran_day, y = .data[[x_axis]])) +  
    geom_jitter(aes(text = paste("\nInfant ID:", baby_id,
                               "\nDay Since Birth:", tran_day,
                               "\n", x_axis, ":", .data[[x_axis]])),   
              width = 0.2, alpha = 0.5, color = "darkblue") +  
    geom_smooth(method = "loess", color = "indianred2", se = FALSE) +  
    labs(
      x = "Days Since Birth to Transfusion",
      y = x_axis,
      title = paste("Population-Level Trend for", x_axis)
    ) + 
    theme_bw() + 
    theme(panel.grid = element_blank())
  
    ggplotly(p, tooltip = "text") %>% 
    layout(
      height = 600)
})

```
